---
title: "QuoteDynamics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{QuoteDynamics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(QuoteDynamics)
```

# QuoteDynamics Package Vignette

## Overview

The `QuoteDynamics` package provides a fast implementation of the Kalman Filter and Smoother Maximum Likelihood Estimation (QuoteDynamics) using the NLopt optimization library. It is designed for advanced statistical analysis in time series and econometrics.

## Installation

### Windows Users
Windows users can directly install the package as it includes the necessary NLopt `.dll` file.

### Linux and macOS Users
Linux and macOS users need to install NLopt separately.

#### Installing NLopt on Linux
You can install NLopt on Linux using the package manager. For example, on Ubuntu:

```bash
sudo apt-get install libnlopt-dev
```

#### Installing NLopt on macOS
On macOS, NLopt can be installed using Homebrew:

```bash
brew install nlopt
```

### Main Function: NLoptKFS

The primary function `NLoptKFS` optimizes parameters for KFS models. 

#### Parameters:
- `start`: Starting values (either a vector or a names list).
- `X`: Data matrix.
- `tau`: Parameter vector.
- `xtol`: Algorithm tolerance.
- `stop_val`: Stopping value.
- `algorithm`: Optimization algorithm (default "LN_NELDERMEAD").
- `hessian`: Whether to compute the Hessian matrix.
- `step_size`: Step size for Hessian computation.
- `verbose`: Verbose output.

### Example

Using the provided datasets `start`, `data`, and `tau`:

```R
Minimum <- QuoteDynamics::quoteDynOptim(start = start, X = data, tau = tau, xtol = 10e-4,
                      stop_val = 10e-10, algorithm = "LN_NELDERMEAD", hessian = TRUE,
                      step_size = 1e-04, verbose = FALSE)
print(Minimum)
```

### Advanced Usage

- `PrintAlgorithms`: This function lists all available optimization algorithms with descriptions. It helps users choose the most suitable algorithm for their specific use case.

### References

This implementation is based on the work of Koopman, S. J., & Durbin, J. (2010), specifically their paper on "A univariate representation of the multivariate Kalman Filter and Smoother". 

### Note

Direct calls to the internal `FastOptim` function are discouraged. Users should interact with the package through `NLoptKFS` for proper error handling and parameter management.

### Included Data

The package comes with sample data sets: `start`, `data`, and `tau`, which are used for demonstration and testing purposes.
